---
title: "FCAT World Cup 2022 Sweep"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tibble)
library(tidyr)
library(dplyr)
library(here)
library(readr)
library(purrr)
library(reactable)

seed = round(16.4 * 1224, digits = 0)
groups <- c("Group A",
            "Group B",
            "Group C",
            "Group D",
            "Group E",
            "Group F",
            "Group G",
            "Group H")

matchdata <- read_csv(here("data", "matchdata.csv"))
payment <- read_csv(here("data", "payment.csv"))

fixtures <- matchdata %>% 
  pivot_wider(names_from = homeaway,
              values_from = c(team, goals, penalties, ycards, rcards, fouls)) %>% 
  mutate(points_hometeam = case_when(goals_hometeam > goals_awayteam ~ 3,
                                     goals_hometeam == goals_awayteam ~ 1,
                                     TRUE ~ 0),
         result_hometeam = case_when(goals_hometeam > goals_awayteam ~ "W",
                                     goals_hometeam == goals_awayteam ~ "D",
                                     TRUE ~ "L"),
         points_awayteam = case_when(goals_awayteam > goals_hometeam ~ 3,
                                     goals_awayteam == goals_hometeam ~ 1,
                                     TRUE ~ 0),
         result_awayteam = case_when(goals_awayteam > goals_hometeam ~ "W",
                                     goals_awayteam == goals_hometeam ~ "D",
                                     TRUE ~ "L"))


hometeam <- fixtures %>% 
            select(matchnumber, roundnumber, date, location, group,
                   team = team_hometeam,
                   result = result_hometeam,
                   goals_for = goals_hometeam,
                   goals_against = goals_awayteam,
                   penalties = penalties_hometeam,
                   ycards = ycards_hometeam,
                   rcards = rcards_hometeam,
                   fouls = fouls_hometeam,
                   points = points_hometeam)

awayteam <- fixtures %>% 
  select(matchnumber, roundnumber, date, location, group,
         team = team_awayteam,
         result = result_awayteam,
         goals_for = goals_awayteam,
         goals_against = goals_hometeam,
         penalties = penalties_awayteam,
         ycards = ycards_awayteam,
         rcards = rcards_awayteam,
         fouls = fouls_awayteam,
         points = points_awayteam)

matchdata_full <- hometeam %>% 
           bind_rows(awayteam)

```

Welcome to the FCAT World Cup 2022 Sweep! Your chance to add a little spice to
your festive season and maybe win a valuable Cash Prize! As usual there are a
range of prizes available to maximise the chances of prolonging your interest.

- **Winner** receives £14
- **Runner up** gets £6

In addition there are some bonus categories with a value of £4 each:

- **Wooden Spoon** for the worst team in the tournament based on results
- **Entertainers** for the team involved in the most goals
- **Thugs** for the dirtiest team

All statistics will be drawn from the official [World Cup
2022](https://www.fifa.com/fifaplus/en/tournaments/mens/worldcup/qatar2022)
website. The organisers reserve the right to tweak the criteria for the bonus
categories, to break ties or generally to keep it sensible. As always in case of
questions over the rules, the organisers decision is _final_.


## The Draw

The draw has been conducted in a rigorously reproducible manner, seeded with a
random number. The random seed used, _`r format(seed, digits = 3)`_, was
generated by multiplying Octobers food price inflation rate (16.4) by the time
the Chancellor finished delivering the Autumn Statement (1224), rounded to an
integer.

Code to reproduce the draw is available
[here](https://github.com/lee269/WC2022).


```{r sweep, echo=FALSE, message=FALSE}


names <-
  c(
    "David Lee", "Emily Dumont", "Maria Prokopiou", "Sarah Thompson",
    "Chris Silwood", "Tom Stone", "Matt Keating", "Michael Archer",
    "Courtney Keane", "Shakti Shah", "Bal Dhoot", "Nick Olney",
    "Stephen Nelson", "Julie Rumsey", "Ros Finney",  "Tim Mordan",
    "Nick Boase", "Sal Burgess", "Helen Hoult", "Nic Feltes",
    "Corinne Wilkins", "Leigh Riley", "Ian Lonsdale", "John Walsh",
    "Clare Gilbert", "Alfie Saunders", "Alexandra Hall", "Michelle Lonsdale",
    "Trevor Hall", "Helen Mason", "Sandra Isted", "Tom Powell"
  )  

countries <- c("Argentina", "Australia", "Belgium", "Brazil", 
               "Cameroon", "Canada", "Costa Rica", "Croatia",
               "Denmark", "Ecuador", "England", "France",
               "Germany", "Ghana", "Iran", "Japan",
               "Mexico", "Morocco", "Netherlands", "Poland",
               "Portugal", "Qatar", "Saudi Arabia", "Senegal",
               "Serbia", "South Korea", "Spain", "Switzerland",
               "Tunisia", "United States", "Uruguay", "Wales")


set.seed(seed)
draw <- tibble(country = sample(countries), name = names) %>% 
  arrange(country)

knitr::kable(draw)

```

## The Prizes {.tabset .tabset-pills}

### Wooden Spoon

This is based on lowest points total, worst goal difference and most goals
conceded, in that sort order.

```{r wooden-spoon, echo=FALSE}

matchdata_full %>% 
  na.omit() %>% 
  filter(group %in% groups) %>%
  group_by(team) %>% 
  summarise(played = n(),
            goals_for = sum(goals_for),
            goals_against = sum(goals_against),
            goals_sum = sum(goals_for) + sum(goals_against),
            goal_diff = sum(goals_for) - sum(goals_against),
            pts = sum(points)) %>% 
  arrange(pts, goal_diff, desc(goals_against)) %>% 
  reactable(pagination = TRUE,
            compact = TRUE,
            columns = list(
              team = colDef(name = "Team"),
              played = colDef(name = "Matches played"),
              goals_for = colDef(name = "Goals for"),
              goals_against = colDef(name = "Goals against"),
              goals_sum = colDef(name = "Goals total"),
              goal_diff = colDef(name = "Goal difference")
            ) )


```

### Entertainers

The entertainment index is the total goals per game plus goals for per game,
with the tie breaker being the number of goals scored (because we all want to
see some buccaneering all out attacking football). 

``` {r entertainers, echo=FALSE}

matchdata_full %>% 
  na.omit() %>% 
  group_by(team) %>% 
  summarise(played = n(),
            goals_for = sum(goals_for),
            goals_against = sum(goals_against),
            goals_sum = sum(goals_for) + sum(goals_against),
            goal_diff = sum(goals_for) - sum(goals_against),
            entertain_index = (goals_sum / played) + (goals_for / played)) %>% 
  arrange(-entertain_index, -goals_for) %>% 
  reactable(pagination = TRUE, 
            compact = TRUE,
            columns = list(
              team = colDef(name = "Team"),
              played = colDef(name = "Matches played"),
              goals_for = colDef(name = "Goals for"),
              goals_against = colDef(name = "Goals against"),
              goals_sum = colDef(name = "Goals total"),
              goal_diff = colDef(name = "Goal difference"),
              entertain_index = colDef(name = "Entertainment Index", format = colFormat(digits = 2))
            ))



```


### Thugs

The 'thug index' is calculated as follows: a red card is 25pts, a yellow 5pts
and a foul 1. These weights are applied to the match data and divided by the
number of matches played.

```{r thugs, echo=FALSE}

matchdata_full %>% 
  na.omit() %>% 
  group_by(team) %>% 
  summarise(played = n(),
            ycards = sum(ycards),
            rcards = sum(rcards),
            fouls = sum(fouls),
            thug_index = ((rcards * 25) + (ycards *5) + fouls) / played) %>% 
  arrange(-thug_index) %>% 
  reactable(pagination = TRUE,
            compact = TRUE,
            columns = list(
              team = colDef(name = "Team"),
              played = colDef(name = "Matches played"),
              ycards = colDef(name = "Yellow cards"),
              rcards = colDef(name = "Red cards"),
              fouls = colDef(name = "Fouls"),
              thug_index = colDef(name = "Thug index", format = colFormat(digits = 2))
            ))

```

## The tournament {.tabset}

### Group stages 


```{r groups, echo=FALSE, message=FALSE, results='asis', warning=FALSE}

league_table <- function(groupname) {
matchdata_full %>% 
  na.omit() %>% 
  filter(group == groupname) %>%
  group_by(team) %>% 
  summarise(played = n(),
            goals_for = sum(goals_for),
            goals_against = sum(goals_against),
            goals_sum = sum(goals_for) + sum(goals_against),
            goal_diff = sum(goals_for) - sum(goals_against),
            ycards = sum(ycards),
            rcards = sum(rcards),
            fouls = sum(fouls),
            thug_index = ((rcards * 25) + (ycards *5) + fouls) / played,
            pts = sum(points)) %>% 
  arrange(-pts, -goal_diff) %>%  
    knitr::kable(caption = paste0("**",groupname,"**"))
}

map(groups, league_table)

```

### Overall standings

```{r overall}

matchdata_full %>% 
  na.omit() %>% 
  group_by(team) %>% 
  summarise(played = n(),
            goals_for = sum(goals_for),
            goals_against = sum(goals_against),
            goals_sum = sum(goals_for) + sum(goals_against),
            goal_diff = sum(goals_for) - sum(goals_against),
            ycards = sum(ycards),
            rcards = sum(rcards),
            fouls = sum(fouls),
            pts = sum(points)) %>% 
  arrange(-pts, -goal_diff) %>% 
  reactable(pagination = FALSE,
            compact = TRUE)



```

### Fixtures
```{r fixtures}

fixtures %>% 
  mutate(fixture = ifelse(is.na(goals_hometeam),
                          paste(team_hometeam, "v", team_awayteam),
                          paste(team_hometeam, goals_hometeam, "-", goals_awayteam, team_awayteam)
                          )
         ) %>% 
  select(matchnumber, date, fixture, location, group, roundnumber) %>% 
  reactable(pagination = FALSE,
            compact = TRUE,
            resizable = TRUE,
            filterable = TRUE,
            columns = list(
            matchnumber = colDef(name = "#"),
            roundnumber = colDef(name = "Round"))
            )

```


## Payment status

Here is the entrants list and payment status. Thanks for your contribution!.
Please correct me if there are any errors.

```{r payment}

payment %>% arrange(Name, desc(Paid)) %>% knitr::kable()


```